---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "**`[[`**Your Name**`]]`**"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Deaths)) + geom_point()
```



```{r first_plot}
ggplot(
  data = covid_data_count,
  mapping = aes(x = state_name, y = Deaths)
) +
  geom_col(width = 0.5, position = position_dodge(0.7)) +  
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Number of deaths related to COVID across various states"
  )
  
```


If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "NY") |> 
  filter(race != "Total") |> 
  ggplot(aes(x = forcats::fct_reorder(race, Deaths), y = Deaths)) + geom_col() + scale_y_continuous(labels = scales::comma) + 
  labs(
    title = "The Number of Deaths related to COVID vs a Specific Race in NY",
    x = "Race",
    y = "Deaths"
  )

```

1. Write code for a plot and describe what you observe?
    Initially, I created a plot to visualize the number of Covid-19 related deaths across different states. Later, my curiosity led me to explore the relationship between deaths and race within a specific city. I chose New York (NY) for this analysis due to its proximity to my hometown, Connecticut (CT).
    
2. Are there any conclusions you can draw?
   Based on the plotted data, it is evident that in New York (NY), the racial group with the highest number of Covid-19 related deaths is White, while the racial group with the lowest number of Covid-19 related deaths is Asian. This observation sheds light on the disparities in Covid-19 outcomes among different racial groups within the city.

3. What were you not able to do due to not having the R knowledge?
   At first, my plot included the category "Total" for a race, and I encountered difficulty in addressing this issue due to my limited knowledge of R. However, upon closer examination of the filtering code that was provided, I managed to resolve the problem by filtering out the data where the race is not categorized as "Total." 
    
4. What other data would be useful to better understand this data?
   To better understand this data it would be useful to have geographic data to analyze the distribution of cases and deaths within New York City. It would also be useful to take a look at vaccination rates among different racial groups to explore the impact of vaccination.




